---
title: "EDA_amy"
author: "Amy Pitts"
date: '2022-04-19'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(tidyverse)
library(gtsummary)
library(mice)
```

## Load Data

```{r}
df = read_excel("Proj 4 data.xlsx") %>%
  janitor::clean_names() %>%
  mutate(
    group = factor(group),
    time_int = case_when(
      time == 1 ~ "baseline",
      time == 2 ~ "6-month",
      time == 3 ~ "12-month"
    ),
    time_int = factor(time_int, levels = c("baseline", "6-month","12-month"))
  )
df %>% head()
```

Data Summary
```{r}
df %>% select(-c(id, time)) %>%
  tbl_summary(by=group) %>%
  add_overall() 


# at baseline 
df %>% select(-id) %>%
  filter(time == 1) %>%
  select(group, sfd) %>%
  tbl_summary(by=group) %>%
  add_overall() %>%
  add_p()
```


Spaghetti plot 

```{r}
df %>%
  mutate(
    group = ifelse(group == 0 , "Control", "Intervention")
  ) %>% 
  ggplot(aes(y=sfd, x=time_int, group=id, color = group))+
  geom_point() +
  geom_line() +
  facet_wrap(~school) +
  labs(
    title = "Symptom free Asthma days by School ",
    y = "Symptom free days in the past two weeks"
  )
```

Why is there is someone that had zero days? 

```{r}
df %>% 
  filter(sfd == 0) %>%
  select(id) %>%
  group_by(id) %>% 
  count() %>%
  arrange(desc(n))
```

So there is one person who didn't have any symptoms in the time span. So that is kinda weird... 

There are 5 people with only 2 zero days. 
```{r}
df %>% 
  filter(id %in% c("2009_1", "3011_1", "3022_1", "4014_1", "4024_1")) %>%
  filter(sfd != 0)
```

Not all on the middle obs but some are.... 


Missing data 

```{r}
data_spread_day = df %>% 
  pivot_wider(names_from = time_int, 
              id_cols = id,
              #names_prefix = "day",
              values_from = sfd
              )

data_spread_day %>%
 select(-c(id)) %>%
  md.pattern(rotate.names = TRUE)
```

So the 3 people that have missing obs for all `sfd` are just going to be excluded. 

As long as the people have baseline obs then we can used them. So 14 people with have observations extrapolated. 










