---
title: "EDA_amy"
author: "Amy Pitts"
date: '2022-04-19'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(tidyverse)
library(gtsummary)
library(mice)
library(lme4)

```

## Load Data

```{r}
df = read_excel("Proj 4 data.xlsx") %>%
  janitor::clean_names() %>%
  mutate(
    group = factor(group),
    time_int = case_when(
      time == 1 ~ "baseline",
      time == 2 ~ "6-month",
      time == 3 ~ "12-month"
    ),
    time_int = factor(time_int, levels = c("baseline", "6-month","12-month")),
    school = factor(school)
  ) %>%
  separate(id, c("id", "one"), sep="_") %>%
  select(-one) %>%
  mutate(id = as.numeric(id))
df %>% head()

```

Data Summary
```{r}
df %>% select(-c(id, time)) %>%
  tbl_summary(by=group) %>%
  add_overall() 


df %>% select(-c(id, time)) %>%
  filter(time_int == "baseline") %>%
  tbl_summary(by=group) %>%
  add_overall() 


# at baseline 
df %>% select(-id) %>%
  filter(time == 1) %>%
  select(group, sfd) %>%
  tbl_summary(by=group) %>%
  add_overall() %>%
  add_p()
```


Spaghetti plot 

```{r}
df %>%
  mutate(
    group = ifelse(group == 0 , "Control", "Intervention")
  ) %>% 
  ggplot(aes(y=sfd, x=time_int, group=id, color = group))+
  geom_point() +
  geom_line() +
  facet_wrap(~school) +
  labs(
    title = "Symptom free Asthma days by School ",
    y = "Symptom free days in the past two weeks"
  )

```

```{r}
df %>%
  mutate(
    group = ifelse(group == 0 , "Control", "Intervention")
  ) %>% 
  ggplot(aes(x=sfd, fill = group))+
  geom_histogram(position = "identity", alpha = 0.7) +
  facet_grid(group~time_int) +
  labs(
    title = "Symptom free Asthma days by School ",
    y = "Count"
  )

```

Want to use sfd_reverse so we are looking at days where there are symptoms. 




Why is there is someone that had zero days? 

```{r}
df %>% 
  filter(sfd == 0) %>%
  select(id) %>%
  group_by(id) %>% 
  count() %>%
  arrange(desc(n))
```

So there is one person who didn't have any symptoms in the time span. So that is kinda weird... 

There are 5 people with only 2 zero days. 
```{r}
df %>% 
  filter(id %in% c("2009", "3011", "3022", "4014", "4024")) %>%
  filter(sfd != 0)
```

Not all on the middle obs but some are.... 


Missing data 

```{r}
data_spread_day = df %>% 
  pivot_wider(
    names_from = time_int, 
    id_cols = id,
    #names_prefix = "day",
    values_from = sfd
)

data_spread_day %>%
 select(-c(id)) %>%
  md.pattern(rotate.names = TRUE)
```

So the 3 people that have missing obs for all `sfd` are just going to be excluded. 

As long as the people have baseline obs then we can used them. So 14 people with have observations extrapolated. 


### Binomial Model 

The Model in question: 

$Y_i \sim Bin(14, p_i)$. \textbf{Need to check assumption of indepencence!}

Let $i$ be subject, $j \in (1,2,3,4)$ be school, and $k \in (1,2,3)$ by the measurement 

\[ 
\log \left(\frac{p_{ijk}}{1-p_{ijk}} \right) = \beta_0 + \beta_1 School_{ij} + \beta_2 Treat_{i} 
+ \beta_3 time_{ijk} + \beta_4 Treat_i \times time_{ijk} + \alpha_{0i} + \alpha_{0j} + \epsilon_{ijk}
\]

Where $\alpha_{0i}$ and $\alpha_{0j}$ are random intercept for subject and school. 


- one model with continuous time 

- one model with categorical time 


[This is a interesting resrouce on negative binomial](https://www.nature.com/articles/s41598-020-73883-7)

## Time as Continous

```{r}
m.p <- glmer(sfd ~  group + time + time * group + (1|id) + (1|school),
              data = df,
              family = poisson)
#summary(m.p)


m.b <- glmer(cbind(sfd, 14-sfd) ~  group + time + time * group + (1|id) + (1|school),
              data = df,
              family = binomial)
#summary(m.b)


m.nb <- glmer.nb(sfd ~ group + time + time * group + (1|id) + (1|school),
              data = df,
              verbose=FALSE)
#summary(m.nb)

## The neg.binomial theta parameter:
#getME(m.nb, "glmer.nb.theta")

var_names = c("intercept", "group1", "time", "group1:time")
mod.nb <- summary(m.nb)
mod.b <- summary(m.b)
mod.p <- summary(m.p)

est_nb_df = mod.nb$coefficients %>% as_tibble() %>%
  janitor::clean_names() %>%
  mutate(var_names = var_names) %>%
  relocate(var_names) %>%
  rename(est_nb = estimate,
         p_nb = pr_z) %>%
  select(var_names, est_nb, p_nb)
est_b_df = mod.b$coefficients %>% as_tibble() %>%
  janitor::clean_names() %>%
  mutate(var_names = var_names) %>%
  relocate(var_names) %>%
  rename(est_b = estimate,
         p_b = pr_z) %>%
  select(var_names, est_b, p_b)
est_p_df = mod.p$coefficients %>% as_tibble() %>%
   janitor::clean_names() %>%
  mutate(var_names = var_names) %>%
  relocate(var_names) %>%
  rename(est_p = estimate,
         p_p = pr_z) %>%
  select(var_names, est_p, p_p)

full_join(est_b_df, est_p_df, by="var_names") %>%
  full_join(est_nb_df, by = "var_names") %>% 
  mutate_at(2:7, round, 3)
```


# Time as Categorical

```{r}
df1 = df %>% mutate(time = factor(time))
m.p <- glmer(sfd ~  group + time + time * group + (1|id) + (1|school),
              data = df1,
              family = poisson)
#summary(m.p)

m.b <- glmer(cbind(sfd, 14-sfd) ~  group + time + time * group + (1|id) + (1|school),
              data = df1,
              family = binomial)
#summary(m.b)


m.nb <- glmer.nb(sfd ~ group + time + time * group + (1|id) + (1|school),
              data = df1,
              verbose=FALSE)

#summary(m.nb)

## The neg.binomial theta parameter:
#getME(m.nb, "glmer.nb.theta")

var_names = c("intercept", "group1", "time2", "time3", "group1:time2","group1:time3")
mod.nb <- summary(m.nb)
mod.b <- summary(m.b)
mod.p <- summary(m.p)

est_nb_df = mod.nb$coefficients %>% as_tibble() %>%
  janitor::clean_names() %>%
  mutate(var_names = var_names) %>%
  relocate(var_names) %>%
  rename(est_nb = estimate,
         p_nb = pr_z) %>%
  select(var_names, est_nb, p_nb)
est_b_df = mod.b$coefficients %>% as_tibble() %>%
  janitor::clean_names() %>%
  mutate(var_names = var_names) %>%
  relocate(var_names) %>%
  rename(est_b = estimate,
         p_b = pr_z) %>%
  select(var_names, est_b, p_b)
est_p_df = mod.p$coefficients %>% as_tibble() %>%
   janitor::clean_names() %>%
  mutate(var_names = var_names) %>%
  relocate(var_names) %>%
  rename(est_p = estimate,
         p_p = pr_z) %>%
  select(var_names, est_p, p_p)

full_join(est_b_df, est_p_df, by="var_names") %>%
  full_join(est_nb_df, by = "var_names")  %>%
  mutate_at(2:7, round, 3)
```


## Transform to Continous outcome 

Section 4.7 (Comparing Ordinal Endpoints by Wilcoxon-Mann-whitney Test) of my book says that if there are 5 or more categories, one can assign a score to each category and treat the change from the baseline of the score as a continous variable. Then we can apply the Equation 4.4. 

Since the ordinal outcome of interest is symptom free days a larger value is better becuase it implies that more of the days were spent without any symptoms. Since larger is better the score I will be assigning to each of the categories is an idenitiy map to the day value. 
```{r}
df_transform <- df %>%
  select(-time_int) %>% 
  pivot_wider(
    names_from = time,
    names_prefix = "obs",
    values_from = sfd 
  ) %>% 
  mutate(
    diff_6month = obs2 - obs1,
    diff_12month = obs3- obs1
  ) %>% 
  rename(baseline = obs1) %>% 
  select(-c(obs2, obs3)) %>%
  pivot_longer(
    c(diff_6month, diff_12month),
    names_to = "time",
    names_prefix = "diff_",
    values_to = "sfd_change" 
  ) %>%
  mutate(
    time = factor(time, levels = c("6month", "12month"))
  )

df_transform %>%
  ggplot(aes(x=time, y=sfd_change, group = id, color= group)) +
  geom_point() +
  geom_line() +
  facet_grid(~school)
```

Now I can fit a regular mixed effect model using this change as a continuous outcome 

```{r}
library("lmerTest")
mod1 <- lmer(sfd_change ~ baseline + group + time + group*time + (1|school)  + (1|id), 
             data = df_transform)

# mod1 <- lmer(sfd_change ~ group + time + group*time + (1|school) + (1|id), 
#              data = df_transform)

summary(mod1)
```

```{r}
#group = ifelse(group == 0 , "Control", "Intervention")
 df %>%
  select(-time_int) %>% 
  pivot_wider(
    names_from = time,
    names_prefix = "obs",
    values_from = sfd 
  ) %>% 
  mutate(
    `Change_6months` = obs2 - obs1,
    `Change_12months` = obs3- obs1
  ) %>% 
  rename(baseline = obs1) %>% 
  select(-c(obs2, obs3)) %>% 
  mutate(
    school = case_when(
      school == 1 ~ "School 1",
      school == 2 ~ "School 2",
      school == 3 ~ "School 3",
      school == 4 ~ "School 4",
    ),
    school = factor( school, 
                     levels = c("School 1", "School 3", "School 2", "School 4"))
  ) %>% 
  select(-c(group,id)) %>% 
  tbl_summary(by=school) %>% 
  add_overall() %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Group Control**") %>%
  modify_spanning_header(c("stat_3", "stat_4") ~ "**Group Intervention**") #%>%


 df %>%
  select(-time_int) %>% 
  pivot_wider(
    names_from = time,
    names_prefix = "obs",
    values_from = sfd 
  ) %>% 
  mutate(
    `Change_6months` = obs2 - obs1,
    `Change_12months` = obs3- obs1
  ) %>% 
  rename(baseline = obs1) %>% 
  select(-c(obs2, obs3)) %>% 
  mutate(
    school = case_when(
      school == 1 ~ "School 1",
      school == 2 ~ "School 2",
      school == 3 ~ "School 3",
      school == 4 ~ "School 4",
    ),
    school = factor( school, levels = c("School 1", "School 3", "School 2", "School 4"))
  ) %>% 
  select(-c(group,id, school)) %>%
  md.pattern(rotate.names = TRUE)
```






### Sample Size Power Stuff 

[PASS all](https://www.ncss.com/software/pass/pass-documentation/)

[Resource:](https://www.ncss.com/wp-content/themes/ncss/pdf/Procedures/PASS/Mixed_Models_Tests_for_Two_Means_in_a_2-Level_Hierarchical_Design-Level-2_Randomization.pdf)


### From the textbook: 

We want to compare Group Intervention vs Group Control. 

Let the observations be denoted 

$$Y_{ij} = 1, \dots, n \ \ (\textrm{main unit}); \ \  j=1, \dots, m_i \ \ (\textrm{subunit})$$

By design we want to set up $m_i= m$ so that all the main units contain the same number of subunits. 

A random effect model is useful to induce the correlation among the subunits as follows: 

$$Y_{ij} = \mu_i + e_{ij}$$

$ \mu_i$ and $e_{ij}$ are independent. 

- The between-cluster variability is: $E[\mu_i] = \mu_y$, $Var(\mu_i) = \sigma_b^2$

- The within-cluster, between-subunit variability is: $E[e_{ij}] = 0$, $Var(e_{ij}) = \sigma^2_w $

Thus the total variance of an observation $Y_{ij}$ and the covariance between observations is $Y_{ij}$ and $Y_{ik}$ within the $i$th cluster are 

$$Var(Y_{ij}) = \sigma^2 = \sigma^2_b + \sigma^2_w$$

$$Cov(Y_{ij}, Y_{ik}) = Var (\mu_i) = \sigma_b^2$$
Thus the correlation among subunits within a cluster equals 
$\rho = \frac{\sigma^2_b}{\sigma^2}$. This correlation is the same across all subunits and is called the compound symmetry model in the repeated measures context. 

We model the data from group $X$ the same way, except that $E[\mu_i]=\mu_x$ for clusters of $X$. 

We can now use the formula used for means for continuous outcomes

$$n= \frac{2\sigma^2 [1+(m-1)\rho](z_{\alpha / 2} + z_{\beta})^2}{m\Delta^2}$$ 

- $m$ we need to find this
- $\beta$ = 1- power
- $\Delta$ mean difference between the two groups and is given by the problem (1/3) 
- $\rho = \frac{\sigma^2_b}{\sigma^2}$
- $\alpha$ need to do a bonforini correction for the 4 observations points 
- $z_{\alpha/2}$, $z_{\beta}$ based on the normal distribution 

```{r}

mod1 <- lmer(sfd_change ~ baseline + group + time + group*time + (1|school)  + (1|id),
             data = df_transform)

mod1 <- lmer(sfd_change ~  group + time + group*time + (1|school)  + (1|id),
             data = df_transform)
# summary(mod1)

# given by the directions 
k = 4 # number of assessments 
n =  30/2 # number of schools 
beta = 0.2 # 1 - power = 1 - 0.8
alpha = 0.05 / (2*k) # bonforinit adjustment 
delta = 1/3 
z.alpha = qnorm(1-alpha)
z.beta = qnorm(1-beta)

# given by the model
m1 = summary(mod1)
sig = m1$varcor %>% as.data.frame() %>% pull(vcov)
sigma.b = sig[2] # between-cluster variability 
sigma.w = sig[1] # within cluster between sub unit variability 
sigma = sigma.b + sigma.w
rho = sigma.b / sigma
rho

# putting it all together to find m 
m = (2*(1-rho)*((z.alpha + z.beta)^2))/(n*(delta^2) - 2*rho*(z.alpha + z.beta)^2)
n = (2*(1+(m-1)*rho)*(z.alpha + z.beta)^2)/(m*(delta^2))
cbind(m, 15*m)


N = (2*(z.alpha+ z.beta)^2/(delta)^2)
m = (1-rho)/(15/N-rho)
N_star = N*(1+(m-1)*rho)
cbind(N_star, m)

```


