---
title: "Yuqi_EDA"
author: "Yuqi Miao ym2771"
date: "4/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(lmerTest)
library(reshape2)
library(nlme)
data = readxl::read_xlsx("Proj 4 data.xlsx") %>% 
  janitor::clean_names() %>% 
  mutate(group = factor(group, levels = c(0,1), labels = c("SOC", "Intervention")),
         school = factor(school))
```


# Data exploration 

## barplot at each time point

verify poisson might be less reasonable
```{r}
data %>% 
  ggplot(aes(x = sfd))+
  geom_bar()+
  facet_grid(time~group)
```

* heavy tail, not appropriate for poisson count

## barplot for 6 v.s. baseline and 12 v.s. baseline 

* 2 goal:
  * see distribution to verify normal trend -- middle high, 2 tail low
  * see if the control/trt group have roughly equan variance to decide 2 sample t test with equal/unequal var (continous) 

```{r}
data_change = data %>% 
  pivot_wider(id_cols = c(id, school,group),
              names_from = time,
              values_from = sfd,
              names_prefix = "time"
              ) %>% 
  mutate(m6_m0 = time2-time1,
         m12_m0 = time3-time1) %>% 
  pivot_longer(c(m6_m0, m12_m0),
               names_to = "compare",
               values_to = "sfd_change") %>% 
  mutate(compare = factor(compare, levels = c("m6_m0", "m12_m0")))

sum_dc = data_change %>% 
  group_by(compare, group) %>% 
  summarize(mean = mean(sfd_change, na.rm = T),
            median = median(sfd_change, na.rm = T))

data_change %>% 
  # filter(compare == "m6_m0") %>% 
  ggplot(aes(x = log(sfd_change), fill = group))+
  # geom_histogram()+
  geom_density(alpha = 0.5)+
  geom_vline(data = sum_dc, aes(xintercept = median,color = group))+
  facet_grid(compare~.)
```

* Truely a normal like distribution, but with right tail
* different vairance: larger variance in the intervention group!

## sphagetti plot 

Goal: Disprove the linear trend and conduct the analysis using time as categorical/ independent measures

```{r}
sp_dc = data_change %>% 
  ggplot(aes(x = compare, y = sfd_change, color = group, group = id))+
  geom_line()+
  geom_point()+
  facet_grid(.~school)

sp_sfd = data %>% 
  ggplot(aes(x = time, y = sfd,color = group, group = id))+
  geom_line()+
  geom_point()+
  stat_summary(aes(group = school),
               geom = "line",
               fun.y = mean, color ="black")+
  facet_grid(.~school)

sp_sfd
ggsave(file = "plots/sphagettie_school_time.png")
```

* by sp_sfd, no clear linear trend along time, not appropriate to treat time as continuous variable


* Box/point plot for intra-class correlation visualization, facet by month. (https://dcricollab.dcri.duke.edu/sites/NIHKR/KR/Intraclass_Correlation_Coefficient_Cheat_Sheet_March_15_2020.pdf)

```{r}
data_change %>% 
  ggplot(aes(x = school, y = sfd_change, color = group))+
  geom_boxplot()+
  facet_grid(.~compare)
```

```{r}
data_change %>% 
  ggplot(aes(x = school, y = time1, color = group))+
  geom_boxplot()
```

* Both btw/within group correlation


# Pilot study analysis


```{r}
# g = lmer(sfd_change~group*compare +(1|school)+(1|id), 
#            data = data_change)
g = lmer(sfd_change~time1+group*compare+(1|id)+(1|school), data = data_change, REML = T)
# 
# data = data %>% 
#   mutate(time = factor(time))
# g = lmer(sfd~time+group+time*group+(1|id)+(1|school), data = data, REML = T)



summary(g)
# texreg::texreg(g)
confint(g)
```

```{r}
summary(g)$coefficients
```

* Conclusion: 
  * When comparing the 6 months and baseline, The increase of sfd in intervention group is 1.67(-0.69, 4.00) more than the increase of sfd in the treatment group
  * When comparing the 12 months and baseline, The increase of sfd in intervention group is 1.40(-0.76 3.63) more than the increase of sfd in the treatment group
  * No significant improvement from intervention group;
  
## estimation of ICC

```{r}
vc = as_tibble(VarCorr(g))
sigma2_2 = vc$vcov[[1]]
sigma2_3 = vc$vcov[[2]]
# sigma2_3=0.4
sigma2_e = vc$vcov[[3]]
```

### if assuming 2 level
```{r}
rho = sigma2_3/(sigma2_2+sigma2_e)
```

### if assuming 3 level

```{r}
rho1 = (sigma2_2+sigma2_3)/(sigma2_2+sigma2_3+sigma2_e)
rho2 = (sigma2_3)/(sigma2_2+sigma2_3+sigma2_e)
# rho2 = 0.05
```



# Study design

## If using continuous outcome and 3 level structure:[1]

$$
y_{ijk} = \beta_0+\delta_{(3)}X_{ijk}+\mu_i+\mu_{j_{i}}+e_{ijk} 
$$

* i for school, j for subjects, k for measures
* $\mu_i \sim N(0, \sigma_3^2)$ random intercept for school
* $\mu_{j_{i}}\sim N(0, \sigma_2^2)$ random intercept for school random intercept for subject
* randomize on school level, $X_{ijk} = X_i = 0/1$ indicating the control/intervention
* Hypothesis: $H_0: \delta_{(3)} = 0$,$H_1: \delta_{(3)} \neq 0$
  * **Q**: $H_1: \delta_{(3)} > 0?$
* calculate N based on normal distribution, with multiple adjustment: $\alpha* = \alpha/4 = 0.05/4 = 0.0125$ for the 4 comparison;
* $\beta = 0.2$

$$
N_2 = \frac{2(1+(\rho_1-\rho_2)N_1-\rho_1)Z_{\alpha, \beta}^2}{N_1N_3\Delta^2-2\rho_2N_1Z_{\alpha, \beta}^2}
$$

```{r}
# alpha_star = 0.05 # multiple test adjust
alpha_star = 0.025/4 # multiple test adjust
# alpha_star = 0.025 # no multiple test adjust
beta = 0.2
z_ab = -qnorm(alpha_star)-qnorm(beta)
delta = 1/3
N3 = 15
# N1 = 4
N1=1
N2 = (2*(1+(rho1-rho2)*N1-rho1)*z_ab^2)/(N1*N3*delta^2-2*rho2*N1*z_ab^2)
N_star = N3*N2
N2
```


### sample size table
```{r}
N2_calc = function(sigma2_3 = vc$vcov[[2]]){
  sigma2_2 = vc$vcov[[1]]
  sigma2_3 = sigma2_3
  sigma2_e = vc$vcov[[3]]
  
  rho1 = (sigma2_2+sigma2_3)/(sigma2_2+sigma2_3+sigma2_e)
  rho2 = (sigma2_3)/(sigma2_2+sigma2_3+sigma2_e)
  
  # alpha_star = 0.025/4 # multiple test adjust
  alpha_star = 0.025 # no multiple test adjust
  beta = 0.2
  z_ab = -qnorm(alpha_star)-qnorm(beta)
  delta = 1/3
  N3 = 15
  N1 = 4
  N2 = (2*(1+(rho1-rho2)*N1-rho1)*z_ab^2)/(N1*N3*delta^2-2*rho2*N1*z_ab^2)
  N_star = N3*N2
  
  return(N2)
}

tibble(sigma2_3 = c(0, 0.3, 0.5)) %>% 
  mutate(N2 = map_dbl(sigma2_3, function(s){
    N2_calc(s)
  }))
```

## If using continuous outcome and 2 level structure: 


A 2 level test structure

$$
y_{ij} = \beta_0+\delta X_{ij}+\mu_i+e_{ij} 
$$

* We regard each time points as a separate point, no within-individual correlation


$$
N_2 = \frac{2(1-\rho)Z^2_{\alpha,\phi}}{N3\Delta^2-2\rho Z^2_{\alpha,\phi}}
$$

```{r}
N2 =(2*(1-rho)*z_ab^2)/(N3*(1/3)^2-2*rho*z_ab^2)
N2
```

* In each treat arm, we need `r N_star` samples, indicating `r m` samples per school

*


# resource

[1]Ahn, C., Heo, M., & Zhang, S. (2014). Sample size calculations for clustered and longitudinal outcomes in clinical research. CRC Press.





